\Large\textbf{General Stage Untangling with Applications to Algorithm Derivation}

### 2014-03-12 paper/paper.tex
split stuff into more files; iterated on proofs

\documentclass{article}
\usepackage{amsmath}
\usepackage{bbm}
\usepackage{stmaryrd}
\usepackage{proof}
\usepackage{amssymb}
\usepackage{mathabx}
\usepackage{cite}
\usepackage{mathpartir}
\usepackage[left=3cm,top=3cm,right=4cm,nohead,bottom=3cm]{geometry}

@@

\input {grammars}

@@

\begin{figure}
\caption{\ellStaged~Static Semantics}
\label{fig:coreStatics}
\begin{mathpar}
\infertypesone [\rmunit] 	{\ty {()}\rmunit}						{\cdot} \and
\infertypesone [int]			{\ty {i}{\rm int}}					{\cdot} \and
\infertypesone [bool]			{\ty {b} \rmbool}						{\cdot} \and
\infertypesone [hyp]			{\ty x A}								{x : A^\bbone \in \Gamma} \and
\infertypestwo [\rmunit]		{\ty {()}{\rmunit}}					{\cdot} \and
\infertypestwo [int]			{\ty {i}{\rm int}}					{\cdot} \and
\infertypestwo [bool]			{\ty {b} \rmbool}						{\cdot} \and
\infertypestwo [hyp]			{\ty x A}								{x : A^\bbtwo \in \Gamma} \and
\infertypesone [let]			{\ty {\letin{x}{e_1}{e_2}}{B}}		{\ty {e_1} A & \ty [\Gamma,x:A^\bbone] {e_2} B} \and
\infertypestwo [let]			{\ty {\letin{x}{e_1}{e_2}}{B}}		{\ty {e_1} A & \ty [\Gamma,x:A^\bbtwo] {e_2} B} \and
\infertypesone [\times I]	{\ty {(e_1,e_2)}{A\times B}}			{\ty {e_1} A & \ty {e_2} B} \and
\infertypesone [\times E_1]	{\ty {\pi_1~e} A}						{\ty e {A\times B}} \and
\infertypesone [\times E_2]	{\ty {\pi_2~e} B}						{\ty e {A\times B}} \and
\infertypestwo [\times I]	{\ty {(e_1,e_2)}{A\times B}}			{\ty {e_1} A & \typestwo {e_2} B} \and
\infertypestwo [\times E_1]	{\ty {\pi_1~e} A}						{\ty e {A\times B}} \and
\infertypestwo [\times E_2]	{\ty {\pi_2~e} B}						{\ty e {A\times B}} \and
\infertypesone [if] 			{\ty {\ifthen{e_1}{e_2}{e_3}} A}	{\ty {e_1} \rmbool & \ty {e_2} A & \ty {e_3} A} \and
\infertypestwo [if]			{\ty {\ifthen{e_1}{e_2}{e_3}} A}	{\ty {e_1} \rmbool & \ty {e_2} A & \ty {e_3} A} \and
\infertypesone [\fut I]		{\ty {\next~e}{\fut A}}				{\typestwo e A} \and
\infertypestwo [\fut E]		{\ty {\prev~e} A}						{\typesone e {\fut A}} \and
\end{mathpar}
\end{figure}

@@

\begin{figure}
\caption{\ellStaged~Erasure Dynamic Semantics}
\label{fig:erasureSemantics}
\begin{mathpar}
\inferreduceone [\rmunit]		{\red {()}{()}}							{\cdot} 											\and
\inferreduceone [int]				{\red {i}{i}}								{\cdot} 											\and
\inferreduceone [bool]			{\red {b}{b}}								{\cdot} 											\and
\inferreduceone [let]				{\red {\letin{x}{e_1}{e_2}}{v_2}}		{\sub [1] & \red{[v_1/x]e_2}{v_2}} 				\and
\inferreducetwo [\rmunit]		{\red {()}{()}}							{\cdot} 											\and
\inferreducetwo [int]				{\red {i}{i}}								{\cdot} 											\and
\inferreducetwo [bool]			{\red {b}{b}}								{\cdot} 											\and
\inferreducetwo [let]				{\red {\letin{x}{e_1}{e_2}}{v_2}}		{\sub [1] & \red{[v_1/x]e_2}{v_2}} 				\and
\inferreduceone [\times I]		{\red {(e_1,e_2)}{\valprod{v_1}{v_2}}}	{\sub [1] & \sub [2]} 							\and
\inferreduceone [\times E_1]		{\red {\pi_1~e}{v_1}}						{\red{e}{\valprod{v_1}{v_2}}} 					\and
\inferreduceone [\times E_2]		{\red {\pi_2~e}{v_2}}						{\red{e}{\valprod{v_1}{v_2}}} 					\and
\inferreducetwo [\times I]		{\red {(e_1,e_2)}{\valprod{v_1}{v_2}}}	{\sub [1] & \sub [2]}								\and
\inferreducetwo [\times E_1]		{\red {\pi_1~e}{v_1}}						{\red{e}{\valprod{v_1}{v_2}}} 					\and
\inferreducetwo [\times E_2]		{\red {\pi_2~e}{v_2}}						{\red{e}{\valprod{v_1}{v_2}}} 					\and
\inferreduceone [if_T] 			{\red {\ifthen{e_1}{e_2}{e_3}}{v}}		{\red{e_1}{\bf true}  & \red{e_2}{v}} 			\and
\inferreduceone [if_F] 			{\red {\ifthen{e_1}{e_2}{e_3}}{v}}		{\red{e_1}{\bf false} & \red{e_3}{v}}			\\
\inferreducetwo [if_T] 			{\red {\ifthen{e_1}{e_2}{e_3}}{v}}		{\red{e_1}{\bf true}  & \red{e_2}{v}} 			\and
\inferreducetwo [if_F] 			{\red {\ifthen{e_1}{e_2}{e_3}}{v}}		{\red{e_1}{\bf false} & \red{e_3}{v}}			\\
\inferreduceone [\fut I]			{\red {\next~e}{\valnext~v}}				{\reducetwosub}									\and
\inferreducetwo [\fut E]			{\red {\prev~e} v}						{\reduceone e {\valnext~v}} 
\end{mathpar}
\end{figure}

\begin{figure}
\caption{\ellStaged~Diagonal Semantics}
\label{fig:refSemantics}
\begin{mathpar}
\inferRedsplitone [\rmunit]		{\red {()}{()}{()}}												{\cdot} 												\and
\inferRedsplitone [int]			{\red {i}{i}{()}}													{\cdot} 												\and
\inferRedsplitone [bool]			{\red {b}{b}{()}}													{\cdot} 												\and
\inferRedsplitone [hyp]			{\red {[u,x]}{u}{x}}												{\cdot}													\and
\inferSpeculation [\rmunit]		{\red {()}{()}}													{\cdot} 												\and
\inferSpeculation [int]			{\red {i}{i}}														{\cdot} 												\and
\inferSpeculation [bool]			{\red {b}{b}}														{\cdot} 												\and
\inferSpeculation [hyp]			{\red {x}{x}}														{\cdot}													\and
\inferRedsplitone [let]			{\red {\letin{x}{e_1}{e_2}}{u_2}{\letin {x}{r_1}{r_2}}}		{\sub [1] & \red{[[u_1,x]/x]e_2}{u_2}{r_2}} 		\and
\inferSpeculation [let]			{\red {\letin{x}{e_1}{e_2}}{\letin{x}{r_1}{r_2}}}				{\sub [1] & \sub [2]}									\and
\inferRedsplitone [\times I]		{\red {(e_1,e_2)}{\valprod{u_1}{u_2}}{\valprod{r_1}{r_2}}}	{\sub [1] & \sub [2]}									\and
\inferRedsplitone [\times E_1]	{\red {\pi_1~e}{u_1}{\pi_1 r}}									{\red {e}{\valprod{u_1}{u_2}}{r}}					\and
\inferRedsplitone [\times E_2]	{\red {\pi_2~e}{u_2}{\pi_2 r}}									{\red {e}{\valprod{u_1}{u_2}}{r}} 					\and
\inferSpeculation [\times I]		{\red {(e_1,e_2)}{(r_1,r_2)}}									{\sub [1] & \sub [2]}									\and
\inferSpeculation [\times E_1]	{\red {\pi_1~e}{\pi_1 r}}										{\red {e}{r}}											\and
\inferSpeculation [\times E_2]	{\red {\pi_2~e}{\pi_2 r}}										{\red {e}{r}} 											\and
\inferRedsplitone [if_T] 		{\red {\ifthen{e_1}{e_2}{e_3}}{u}{(r_1;r_2)}}					{\red{e_1}{\bf true }{r_1} & \red{e_2}{u}{r_2}}	\and
\inferRedsplitone [if_F] 		{\red {\ifthen{e_1}{e_2}{e_3}}{u}{(r_1;r_3)}}					{\red{e_1}{\bf false}{r_1} & \red{e_3}{u}{r_3}}	\and
\inferSpeculation [if] 			{\red {\ifthen{e_1}{e_2}{e_3}}{\ifthen{r_1}{r_2}{r_3}}}		{\sub [1] & \sub [2] & \sub [3]}					\\
\inferRedsplitone [\fut I]		{\red {\next~e}{()}{r}}											{\rstwosub}											\and
\inferSpeculation [\fut E]		{\red {\prev~e}{r}}												{\rsone {e}{()}{r}}
\end{mathpar}
\end{figure}

\subsection{Splitting}

@@

\begin{figure*}
\caption{Type Splitting}
\label{fig:typeSplit}
\begin{mathpar}
\infer [\mathrm{unit}\tsplito]	{\rmunit \tsplito [\rmunit,\rmunit]}{\cdot} \and
\infer [\mathrm{int}\tsplito]	{{\rm int} \tsplito [{\rm int}, \rmunit]}{\cdot} \and
\infer [\mathrm{bool}\tsplito]	{{\rm bool} \tsplito [{\rm bool}, \rmunit]}{\cdot} \and
\infer [\times\tsplito]
	{A\times B \tsplito [A_1 \times B_1, A_2 \times B_2]}
	{A \tsplito [A_1, A_2] & B \tsplito [B_1,B_2]} \and
\infer [\fut\tsplito]
	{\fut~A \tsplito [{\rm unit}, A']}
	{A \tsplits A'} \and
\end{mathpar}
\end{figure*}

\begin{figure*}
\caption{Value Splitting}
\label{fig:valSplit}
\begin{mathpar}
\infer [\rmunit\vsplito]		{() \vsplito [(), ()]}{\cdot} \and
\infer [\mathrm{int}\vsplito]	{i \vsplito [i,()]}{\cdot} \and
\infer [\mathrm{bool}\vsplito]	{b \vsplito [b,()]}{\cdot} \and
\infer [\times\vsplito]	
	{\valprod{v_1}{v_2} \vsplito [\valprod{u_1}{u_2}, \valprod{w_1}{w_2}]}
	{v_1 \vsplito [u_1,w_1] & v_2 \vsplito [u_2,w_2]} \and
\infer [\fut\vsplito]
	{\valnext~v \vsplito [(),v']}
	{v \vsplits v'} 
\end{mathpar}
\end{figure*}

\begin{figure*}
\caption{Context Splitting}
\label{fig:contSplit}
\begin{mathpar}
\infer {\emptyC \csplit [\emptyC~;~\emptyC]}{\cdot} \and
\infer {\Gamma, v : A^\bbone \csplit [\Gamma_1,v:A_1; \Gamma_2,v:A_2]}{\Gamma \csplit [\Gamma_1; \Gamma_2] & A \tsplito [A_1, A_2]} \and
\infer {\Gamma, v : A^\bbtwo \csplit [\Gamma_1; \Gamma_2,v:A']}{\Gamma \csplit [\Gamma_1; \Gamma_2] & A \tsplits A'} 
\end{mathpar}
\end{figure*}

\begin{figure*}
\caption{Term Splitting}
\label{fig:termSplit}
\begin{mathpar}
\infersplitone [unit]			{\spl {()}{\rmunit}{((),()),\_.()}}															{\cdot} \and
\infersplittwo [unit]			{\spl {()}{\rmunit}{(),\_.()}}																{\cdot} \and
\infersplitone [int]			{\spl {i}{\rm int}{(i,()),\_.()}}															{\cdot} \and
\infersplittwo [int]			{\spl {i}{\rm int}{(),\_.i}}																	{\cdot} \and
\infersplitone [\times I] 	{\spl {(e_1,e_2)}{A\times B}
									{ \begin{array}{l}
									\left(
										\letin{y_1}{c_1}{\letin{y_2}{c_2}{
										((\pi_1 y_1, \pi_1 y_2),(\pi_2 y_1, \pi_2 y_2))}}
									\right), \\
									l.(\projbind 1,\projbind 2)
									\end{array}}}																				{\sub 1 A & \sub 2 B} \and
\infersplittwo [\times I]	{\spl {(e_1,e_2)}{A\times B}{(p_1,p_2),l.(\projbind 1,\projbind 2)}}					{\sub 1 A & \sub 2 B} \and
\infersplitone [\times E_1]	{\spl {\pi_1~e}{A}{\letin{y}{c}{(\pi_1(\pi_1 y),\pi_2 y)},l.\pi_1~r}}					{\sub {} {A\times B}} \and
\infersplitone [\times E_2]	{\spl {\pi_2~e}{B}{\letin{y}{c}{(\pi_2(\pi_1 y),\pi_2 y)},l.\pi_2~r}}					{\sub {} {A\times B}} \and
\infersplittwo [\times E_1]	{\spl {\pi_1~e}{A}{p,l.\pi_1~r}}															{\sub {} {A\times B}} \and
\infersplittwo [\times E_2]	{\spl {\pi_2~e}{B}{p,l.\pi_2~r}}															{\sub {} {A\times B}} \and
\infersplitone [hyp]			{\spl {v}{A}{(x,()),\_.()}}																	{x:A^\bbone \in \Gamma} \and
\infersplittwo [hyp]			{\spl {v}{A}{(),\_.x}}																		{x:A^\bbtwo \in \Gamma} \\
\infersplitone [let]			{\spl {\letin{x}{e_1}{e_2}}{B}{
								\begin{array}{l}
									\left(\talllet{y_1}{c_1}{
									\talllet{y_2}{(\letin{x}{\pi_1 y_1}{c_2})}{
									(\pi_1 y_2, (\pi_2 y_1, \pi_2 y_2))}}\right), \\
									l.\letin{x}{(\projbind 1)}{\projbind 2}
								\end{array}}}																					{\sub 1 A & \sub [\Gamma,x:A^\bbone] 2 B} \and
\infersplittwo [let]			{\spl {\letin{x}{e_1}{e_2}}{B}{(p_1,p_2), l.\letin{x}{(\projbind 1)}{\projbind 2}}}	{\sub 1 A & \sub [\Gamma,x:A^\bbtwo] 2 B} \and
\infersplitone [\fut I]		{\spl {\next~e}{\fut A}{((),p),l.r}}														{\splittwosub {} A} \and
\infersplittwo [\fut E]		{\spl {\prev~e}{A} {\pi_2~c,l.r}}															{\splitonesub {} {\fut A}} \and
\infersplitone [if]			{\spl {\left( \tallif {e_1}{e_2}{e_3} \right)}{A} 
								{\begin{array}{l}
									\left(\talllet{z}{c_1}{\tallif {\pi_1 z}
										{\letin{y}{c_2}{(\pi_1 y, (\pi_2 z, \inl(\pi_2 y)))}}
										{\letin{y}{c_3}{(\pi_1 y, (\pi_2 z, \inr(\pi_2 y)))}}}
									\right), \\
									l.(\projbind 1;\caseof{\pi_2 l}{l_2.r_2}{l_3.r_3})
								\end{array}}}																					{\sub 1 \rmbool & \sub 2 A & \sub 3 A } \and
\infersplittwo [if]			{\spl {\left(\tallif {e_1}{e_2}{e_3}\right)}{A} 
								{(p_1,p_2,p_3),l.\left(\tallif{(\projbind 1)}{(\projbind 2)}{(\projbind 3)}\right)}}	{\sub 1 \rmbool & \sub 2 A & \sub 3 A } \and
\end{mathpar}
\end{figure*}

@@

\begin{itemize}
\item {\bf Int.} Suppose that
\begin{enumerate}
\item $\splitone{i}{\rm int}{(i,()),\_.()}$
\item $\rsone{i}{v}{q}$
\item $\reducetwo{r}{w}$
\end{enumerate}
By inversion of reduction relations, we know
\begin{enumerate}
\item $v$ has the form $i$
\item $q$ has the form $()$
\item $w$ has the form $()$
\end{enumerate}
It is directly provable that,
\begin{enumerate}
\item $\reduce{(i,())}{\valprod{i}{()}}$
\item $\reduce{[()/l]()}{()}$
\end{enumerate}
Which is what we need to show.

\item {\bf Tuple.} Suppose that
\begin{enumerate}
\item $\splitone {(e_1,e_2)}{A\times B}
		{ \begin{array}{l}
		(\letin{y_1}{c_1}{\letin{y_2}{c_2}{((\pi_1 y_1, \pi_1 y_2),(\pi_2 y_1, \pi_2 y_2))}}), \\
		l.(\letin{l_1}{\pi_1 l}{r_1},\letin{l_2}{\pi_2 l}{r_2})
		\end{array}} $
\item $\rsone{(e_1,e_2)}{v}{q}$
\item $\reducetwo{q}{w}$
\end{enumerate}
By inversion of the splitting and reduction rules, we know
\begin{enumerate}
\item $\splitone{e_1}{A}{c_1,l_1.r_1}$ and $\splitone{e_2}{B}{c_2,l_2.r_2}$
\item $v$ has the form $\valprod{v_1}{v_2}$
\item $q$ has the form $(q_1,q_2)$
\item $\rsone{e_1}{v_1}{q_1}$ and $\rsone{e_2}{v_2}{q_2}$
\item $w$ has the form $\valprod{w_1}{w_2}$
\item $\reducetwo{q_1}{w_1}$ and $\reducetwo{q_2}{w_2}$
\end{enumerate}
This gives us enough to invoke the inductive hypothesis for the each side, yielding
\begin{enumerate}
\item $\reduce{c_1}{\valprod{v_1}{u_1}}$ 
\item $\reduce{[u_1/l_1]r_1}{w_1}$ 
\item $\reduce{c_2}{\valprod{v_2}{u_2}}$ 
\item $\reduce{[u_2/l_2]r_2}{w_2}$ 
\end{enumerate}
It is now directly provable from the evaluation rules that
\begin{enumerate}
\item $\reduce{\letin{y_1}{c_1}{\letin{y_2}{c_2}{((\pi_1 y_1, \pi_1 y_2),(\pi_2 y_1, \pi_2 y_2))}}}{\valprod {\valprod{v_1}{v_2}}{\valprod{u_1}{u_2}}}$
\item $\reduce{[\valprod{u_1}{u_2}/l](\letin{l_1}{\pi_1 l}{r_1},\letin{l_2}{\pi_2 l}{r_2})}{\valprod{w_1}{w_2}}$
\end{enumerate}
Which is what we wanted to show.

\item {\bf Projection.} Suppose that
\begin{enumerate}
\item $\splitone {\pi_1~e}{A}{\letin{y}{c}{(\pi_1(\pi_1 y),\pi_2 y)},l.\pi_1~r}$
\item $\rsone{\pi_1~e}{v}{q}$
\item $\reducetwo{q}{w}$
\end{enumerate}
By inversion of the splitting and reduction rules, we have
\begin{enumerate}
\item $\splitone {e}{A \times B}{c,l.r}$
\item $q$ has the form $\pi_1~q'$
\item $\rsone{e}{\valprod{v}{v_2}}{q'}$
\item $\reducetwo{q'}{\valprod{w}{w_2}}$
\end{enumerate}
This is enough to invoke our inductive hypothesis, yielding
\begin{enumerate}
\item $\reduce{c}{\valprod{\valprod{v}{v_2}}{u}}$
\item $\reduce{[u/l]r}{\valprod{w}{w_2}}$
\end{enumerate}
From this, it can be directly proven that
\begin{enumerate}
\item $\reduce{\letin{y}{c}{(\pi_1(\pi_1 y),\pi_2 y)}}{\valprod{v}{u}}$
\item $\reduce{[u/l](\pi_1~r)}{w}$
\end{enumerate}
Which is what we wanted to show.

\item {\bf If.} Suppose that
\begin{enumerate}
\item $\splitone {\left( \tallif {e_1}{e_2}{e_3} \right)}{A} 
								{\begin{array}{l}
									\left(\tallif {\pi_1c_1}
										{\letin{y}{c_2}{(\pi_1 y, \inl(\pi_2 y))}}
										{\letin{y}{c_3}{(\pi_1 y, \inr(\pi_2 y))}}
									\right), \\
									l.\left(\caseof{l}{l_2.r_2}{l_3.r_3}
									\right)
								\end{array}}$
\item $\rsone{\ifthen {e_1}{e_2}{e_3}}{v}{q}$
\item $\reducetwo{q}{w}$
\end{enumerate}
By inversion of the splitting rule, we have
\begin{enumerate}
\item $\splitonesub 1 \rmbool$ and $\splitonesub 2 A$ and $\splitonesub 3 A$
\end{enumerate}
There are two rules for evaluation of an if.  We proceed by inverting the if-true and evaluation rules first:
\begin{enumerate}
\item $q$ has the form $(q_1;q_2)$
\item $\reducetwo{q_1}{w_1}$ and $\reducetwo{q_2}{w}$
\item $\rsone{e_1}{\bf true}{q_1}$ and $\rsone{e_2}{v}{q_2}$
\end{enumerate}
We can now invoke the inductive hypothesis for the predicate to yield
\begin{enumerate}
\item $\reduce{c_1}{\valprod{\bf true}{u_1}}$
\item $\reduce{[u_1/l_1]q_1}{w_1}$
\end{enumerate}
We can also invoke the inductive hypothesis for the branch to yield
\begin{enumerate}
\item $\reduce{c_2}{\valprod{v}{u_2}}$
\item $\reduce{[u_2/l_2]r_2}{w}$
\end{enumerate}
From this, it can be directly proven that
\begin{enumerate}
\item $\reduce{
		\left(\tallif {\pi_1c_1}
			{\letin{y}{c_2}{(\pi_1 y, \inl(\pi_2 y))}}
			{\letin{y}{c_3}{(\pi_1 y, \inr(\pi_2 y))}}
		\right)}{\valprod{v}{\inl~u_2}}$
\item $\reduce{[\inl~u_2/l](\caseof{l}{l_2.r_2}{l_3.r_3})}{w}$
\end{enumerate}
Everything works similarly for the if-false case.  Either way, we have what we want to show.

@@

\begin{figure*}
\caption{Sum and Function Typing}
\label{fig:advancedTypes}
\begin{mathpar}
\infertypesone [\to I]	{\ty {\lam{x}{A}{e}} {A \to B}}				{\ty [\Gamma,x:A^\bbone] e B} 																\and
\infertypesone [\to E]	{\ty {e_1~e_2} {B}}							{\ty {e_1} {A \to B} & \ty {e_2} A} 														\and
\infertypestwo [\to I]	{\ty {\lam{x}{A}{e}} {A \to B}}				{\ty [\Gamma,x:A^\bbtwo] e B} 																\and
\infertypestwo [\to E]	{\ty {e_1~e_2} {B}}							{\ty {e_1} {A \to B} & \ty {e_2} A} 														\and
\infertypesone [+ I_1]	{\ty {\inl~e} {A + B}}						{\ty e A} 																						\and
\infertypesone [+ I_2]	{\ty {\inr~e} {A + B}}						{\ty e B} 																						\\
\infertypestwo [+ I_1]	{\ty {\inl~e} {A + B}}						{\ty e A} 																						\and
\infertypestwo [+ I_2]	{\ty {\inr~e} {A + B}}						{\ty e B} 																						\and
\infertypesone [+ E]		{\ty {\caseof{e_1}{x_2.e_2}{x_3.e_3}} C}	{\ty {e_1}{A+B} & \ty[\Gamma,x_2:A^\bbone]{e_2} C & \ty[\Gamma,x_3:A^\bbone]{e_3} C} 	\and
\infertypestwo [+ E]		{\ty {\caseof{e_1}{x_2.e_2}{x_3.e_3}} C}	{\ty {e_1}{A+B} & \ty[\Gamma,x_2:A^\bbtwo]{e_2} C & \ty[\Gamma,x_3:A^\bbtwo]{e_3} C} 	\and
\end{mathpar}
\end{figure*}

\begin{figure*}
\caption{Sum and Function Erasure Evaluation}
\label{fig:advancedTypes}
\begin{mathpar}
\inferreduceone [\to I]	{\red {\lam{x}{A}{e}} {\vallam{x}{M}}}			{\cdot} 																		\and
\inferreduceone [\to E]	{\red {e_1~e_2} {v'}}								{\red {e_1} {\vallam{x}{M}} & \red{e_2}v & \red {[v/x]M} {v'}}			\and
\inferreducetwo [\to I]	{\red {\lam{x}{A}{e}} {\vallam{x}{M}}}			{\cdot} 																		\and
\inferreducetwo [\to E]	{\red {e_1~e_2} {v'}}								{\red {e_1} {\vallam{x}{M}} & \red{e_2}v & \red {[v/x]M} {v'}}			\and
\inferreduceone [+ I_1]	{\red {\inl~e} {\valinl~v}}						{\sub} 																			\and
\inferreduceone [+ I_2]	{\red {\inr~e} {\valinr~v}}						{\sub} 																			\\
\inferreducetwo [+ I_1]	{\red {\inl~e} {\valinl~v}}						{\sub} 																			\and
\inferreducetwo [+ I_2]	{\red {\inr~e} {\valinr~v}}						{\sub} 																			\\
\inferreduceone [+ E_1]	{\red {\caseof{e_1}{x_2.e_2}{x_3.e_3}}{v'}}	{\red {e_1}{\valinl~v} & \red {[v/x_2]e_2}{v'}} 							\and
\inferreduceone [+ E_2]	{\red {\caseof{e_1}{x_2.e_2}{x_3.e_3}}{v'}}	{\red {e_1}{\valinr~v} & \red {[v/x_3]e_3}{v'}} 							\and
\inferreducetwo [+ E_1]	{\red {\caseof{e_1}{x_2.e_2}{x_3.e_3}}{v'}}	{\red {e_1}{\valinl~v} & \red {[v/x_2]e_2}{v'}} 							\and
\inferreducetwo [+ E_2]	{\red {\caseof{e_1}{x_2.e_2}{x_3.e_3}}{v'}}	{\red {e_1}{\valinr~v} & \red {[v/x_3]e_3}{v'}} 							\and
\end{mathpar}
\end{figure*}

\begin{figure*}
\caption{Sum and Function Diagonal Evaluation}
\label{fig:advancedTypes}
\begin{mathpar}
\inferRedsplitone [\to I]	{\red {\lam{x}{A}{e}}{\vallam{x}{e}}{()}}			{\sub} 																	\and
\inferSpeculation [\to I]	{\red {\lam{x}{A}{e}}{\vallam{x}{r}}}				{\sub} 																	\and
\inferRedsplitone [\to E]	{\red {e_1~e_2}{v'}{(r_1;\letin{x}{r_2}{r'})}}		{\red {e_1}{\vallam{x}{M}}{r_1} 
																							& \sub [2] & \red{[[e,x]/x]M}{v'}{r'}} 							\and
\inferSpeculation [\to E]	{\red {e_1~e_2}{r_1~r_2}}							{\sub [1] & \sub[2]}													\and
\inferRedsplitone [+ I_1]	{\red {\inl~e} {\valinl~v}{r}}						{\sub} 																	\and
\inferSpeculation [+ I_1]	{\red {\inl~e} {\valinl~r}}							{\sub} 																	\\
\inferRedsplitone [+ I_2]	{\red {\inr~e} {\valinr~v}{r}}						{\sub} 																	\and
\inferSpeculation [+ I_2]	{\red {\inr~e} {\valinr~r}}							{\sub} 																	\\
\inferRedsplitone [+ E_1]	{\red {\caseof{e_1}{x_2.e_2}{x_3.e_3}}{v'}
								{\letin{x_2}{r_1}{r_2}}}								{\red {e_1}{\valinl~v}{r_1} & \red {[[v,x_2]/x_2]e_2}{v'}{r_2}} \and
\inferRedsplitone [+ E_1]	{\red {\caseof{e_1}{x_2.e_2}{x_3.e_3}}{v'}
								{\letin{x_3}{r_1}{r_3}}}								{\red {e_1}{\valinr~v}{r_1} & \red {[[v,x_3]/x_3]e_3}{v'}{r_3}} \and
\inferSpeculation [+ E]		{\red {\caseof{e_1}{x_2.e_2}{x_3.e_3}}
								{\caseof{r_1}{x_2.r_2}{x_3.r_3}}}				{\sub [1] & \sub [2] & \sub[3]}					 						\and
\end{mathpar}
\end{figure*}

\begin{figure*}
\caption{Sum and Function Splitting}
\label{fig:advancedTypes}
\begin{mathpar}
\infersplitone	 [\to I]	{\spl {\lam{x}{A}{e}}{A \to B}{\vallam{x}{c},
							\_.(\vallam{y}{\letin{x}{\pi_1 y}{\letin{l}{\pi_2 y}{r}}})}}		{\sub [\Gamma,x:A^\bbone] {} B} 						\and
\infersplittwo [\to I]	{\spl {\lam{x}{A}{e}}{A \to B}{p,l.(\vallam{x}{M})}}				{\sub [\Gamma,x:A^\bbtwo] {} B} 						\and
\infersplitone [\to E]	{\spl {e_1~e_2}{B}{\ldots}}											{\sub 1 {A \to B} & \sub 2 A}							\and
\infersplittwo [\to E]	{\spl {e_1~e_2}{B}{(p_1,p_2),l.(\projbind 1)(\projbind 2)}}		{\sub 1 {A \to B} & \sub 2 A}							\and
\infersplitone [+ I_1]	{\spl {\inl~e}{A+B}{\letin{y}{c}{(\inl(\pi_1y),\pi_2 y)},l.r}}	{\sub {} A}												\and
\infersplitone [+ I_2]	{\spl {\inr~e}{A+B}{\letin{y}{c}{(\inr(\pi_1y),\pi_2 y)},l.r}}	{\sub {} A}												\and
\infersplittwo [+ I_1]	{\spl {\inl~e}{A+B}{l.\inl~r}}										{\sub {} A}												\and
\infersplittwo [+ I_1]	{\spl {\inr~e}{A+B}{l.\inr~r}}										{\sub {} A}												\and
\infersplitone [+ E]		{\spl {\left(\tallcase{e_1}{x_2.e_2}{x_3.e_3}\right)}{C}
								{\begin{array}{l}\left(
								\talllet{y_1}{c_1}{
									\tallcase{\pi_1y_1}
									{x_2.\letin{y_2}{c_2}{(\pi_1y_2,(\pi_2y_1,\inl(\pi_2y_2)))}}
									{x_3.\letin{y_3}{c_3}{(\pi_1y_3,(\pi_2y_1,\inr(\pi_2y_3)))}}
								}\right), \\
								l.\left(\talllet{z}{(\projbind 1)}{
									\tallcase{\pi_2 l}
									{l_2.\letin{x_2}{z}{r_2}}
									{l_3.\letin{x_3}{z}{r_3}}}\right)\end{array}}}				{\sub 1 {A+B} & \sub [\Gamma,x_2:A^\bbone] 2 C 
																									& \sub [\Gamma,x_3:A^\bbone] 3 C} 						\and
\infersplittwo [+ E]		{\spl {\left(\tallcase{e_1}{x_2.e_2}{x_3.e_3}\right)}{C}
								{(p_1,p_2,p_3), l.\left(\tallcase{(\projbind 1)}
									{x_2.\projbind 2}{x_3.\projbind 3}\right)}}				{\sub 1 {A+B} & \sub [\Gamma,x_2:A^\bbtwo] 2 C 
																									& \sub [\Gamma,x_3:A^\bbtwo] 3 C} 						\and
\end{mathpar}
\end{figure*}

@@

>>>>

\documentclass{article}
\usepackage{amsmath}
\usepackage{bbm}
\usepackage{stmaryrd}
\usepackage{proof}
\usepackage{amssymb}
\usepackage{mathabx}
\usepackage{float}
\usepackage{cite}
\usepackage{mathpartir}
\usepackage[left=3cm,top=3cm,right=4cm,nohead,bottom=3cm]{geometry}

@@

\input {grammars}
\input {semantics}
\input {splitting}

@@

@@

\section{The Stage-Splitting Transformation}

@@

@@

@@

@@

\section{Appendix}

\subsection {Equivalence of Diagonal Evaluation and Splitting (for \ellStaged)}

\begin{itemize}
% INT, BOOL, UNIT
\item {\bf Int.} Suppose that
\begin{enumerate}
\item $\splitone{i}{\rm int}{(i,()),\_.()}$
\item $\rsone{i}{v}{q}$
\item $\reducetwo{q}{w}$
\end{enumerate}
By inversion of reduction relations, we know
\begin{enumerate}
\item $v$ has the form $i$
\item $q$ has the form $()$
\item $w$ has the form $()$
\end{enumerate}
It is directly provable that,
\begin{enumerate}
\item $\reduce{(i,())}{\valprod{i}{()}}$
\item $\reduce{[()/\_]()}{()}$
\end{enumerate}
Which is what we need to show.

\item {\bf Boolean.} Same as int.
\item {\bf Unit.} Same as int.

% TUPLE
\item {\bf Tuple.} Suppose that
\begin{enumerate}
\item $\splitone {(e_1,e_2)}{A\times B}
		{ \begin{array}{l}
		(\letin{y_1}{c_1}{\letin{y_2}{c_2}{((\pi_1 y_1, \pi_1 y_2),(\pi_2 y_1, \pi_2 y_2))}}), \\
		l.(\letin{l_1}{\pi_1 l}{r_1},\letin{l_2}{\pi_2 l}{r_2})
		\end{array}} $
\item $\rsone{(e_1,e_2)}{v}{q}$
\item $\reducetwo{q}{w}$
\end{enumerate}
By inversion of the splitting and reduction rules, we know
\begin{enumerate}
\item $\splitone{e_1}{A}{c_1,l_1.r_1}$ and $\splitone{e_2}{B}{c_2,l_2.r_2}$
\item $v$ has the form $\valprod{v_1}{v_2}$
\item $q$ has the form $(q_1,q_2)$
\item $\rsone{e_1}{v_1}{q_1}$ and $\rsone{e_2}{v_2}{q_2}$
\item $w$ has the form $\valprod{w_1}{w_2}$
\item $\reducetwo{q_1}{w_1}$ and $\reducetwo{q_2}{w_2}$
\end{enumerate}
This gives us enough to invoke the inductive hypothesis for the each side, yielding
\begin{enumerate}
\item $\reduce{c_1}{\valprod{v_1}{u_1}}$ 
\item $\reduce{[u_1/l_1]r_1}{w_1}$ 
\item $\reduce{c_2}{\valprod{v_2}{u_2}}$ 
\item $\reduce{[u_2/l_2]r_2}{w_2}$ 
\end{enumerate}
It is now directly provable from the evaluation rules that
\begin{enumerate}
\item $\reduce{\letin{y_1}{c_1}{\letin{y_2}{c_2}{((\pi_1 y_1, \pi_1 y_2),(\pi_2 y_1, \pi_2 y_2))}}}{\valprod {\valprod{v_1}{v_2}}{\valprod{u_1}{u_2}}}$
\item $\reduce{[\valprod{u_1}{u_2}/l](\letin{l_1}{\pi_1 l}{r_1},\letin{l_2}{\pi_2 l}{r_2})}{\valprod{w_1}{w_2}}$
\end{enumerate}
Which is what we wanted to show.
